% -*- mode: latex; tex-main-file: "thesis.tex"; -*-

%{\setbox0\hbox{\vtop{%
%\begin{flushright}%
%    \def\baselinestretch{1}\slshape Pithy quote.\\
%\vskip 1em
%- Source of quote
%\end{flushright}%
%\null}}\ht0=0pt\dp0=0pt\box0}



%\qack{Look at these grand men. Which of you wouldn't consider it the
%highlight of his career just to associate with them for even one day?
%\vskip 1em - Lou Gehrig
%}

%\qack{\textit{The crowd makes the ballgame.}
%\vskip 1em - Ty Cobb
%}

%\vspace*{100\p@}%
%{\hfill {\sfbHuge{Acknowledgments}}}
%\vskip 40\p@

\chapter*{Acknowledgments}

{\setbox0\hbox{\vtop{%
\begin{flushleft}%
\vskip -2.5in
    \def\baselinestretch{1}
{\slshape \textit{The crowd makes the ballgame.} 
\vskip 0.1em - Ty Cobb}
\end{flushleft}%
\null}}\ht0=0pt\dp0=0pt\box0}


\addcontentsline{toc}{chapter}{Acknowledgments}

\noindent
My advisor, Hari Balakrishnan, once said that one secret to success is
to surround yourself with people who are smarter than you are.  This
task must be incredibly difficult for him.  I am extremely fortunate to
have crossed paths with Hari and to have had the opportunity to work
with him so closely for many years.  Hari has guided me with amazing
expertise.  His sharp intellect and ability to quickly grasp the
important and interesting aspects of a problem are matched only by his
fantastic taste in research problems.  Hari opened (and closed) all of
the right doors for me; he gave me the opportunity to succeed at every
corner, and he continually encouraged me to consider the broader impact of my
research while allowing me the freedom to pursue problems I found most
interesting.

Jennifer Rexford has been another fantastic mentor.  She shares my
insatiable passion for details and practical problems.  Her patience,
willingness to consider new ideas, and ability to explain complicated
concepts make her a fun person to work with and should serve as a model
for every advisor.  Jennifer introduced me to Internet routing, and many
of her papers provided inspiration for several chapters in this
dissertation.  She has also been a teacher, a useful sounding board, and
a continual source of sound advice.

I owe a great debt to the other two members of my committee, David Clark
and Frans Kaashoek.  David read this dissertation with vigilance, and my
discussions with him, both during the writing process and throughout my
graduate career, helped me strengthen many of the conceptual aspects of
this dissertation.  Frans provided invaluable guidance in clarifying the
presentation of many of the ideas in this dissertation.  His enthusiasm for
both the work in this dissertation and other projects we worked on
together have kept 
me optimistic and excited about my research.  When I'd think
some of my projects were fruitless or boring, Frans 
always provided the boost I needed.  On the squash court, he continually
reminds 
me about the virtues of humility.

I am lucky to have Ramesh Johari and David Andersen as colleagues.  I've
had great fun working with both of them and learning from their
different research styles.  Ramesh's penchant for precision---both in
math and in English---and his tenacity in problem solving make him a
pleasure to work with.  I have great respect for him as
a researcher, a mentor, and a friend.  Dave's energy, fun-loving nature,
and willingness to put up with my continual questions while always
questioning my own research was one of the most
valuable assets throughout my graduate career.  Dave was my officemate
and housemate, but he is also a great collaborator and a terrific
friend.  Some of my best times in graduate school were spent cycling
through Massachusetts and talking research 
with Dave.

One of the great things about MIT is that you don't have to try very
hard to surround yourself with people who are smarter than you are.  My
research and career benefited from the influence of many faculty
members, including Rod Brooks, John Guttag, David Karger, Daniel
Jackson, Dina Katabi, Barbara Liskov, Sam Madden, Robert Morris, and Ron
Rivest.  It
has been a joy to share my graduate career with Magdalena Balazinska;
she is a tremendous colleague and friend. Thanks to her, I remembered
to eat lunch most of the time.  I have learned a lot
from exchanging ideas with Alex Snoeren, who was also a tremendous
support during my job search.
Both Michel  
Goraczko and Dorothy Curtis 
were helpful and patient with my technical questions and
emergencies.  I have also had fantastic officemates.  Jaeyeon Jung and
Mythili Vutukuru have been fun to work with; I hope we continue
working together.  Michael Walfish's clarity of expression---not to
mention his enthusiasm for a good argument---made G982 an exciting place
to work.  Discussions with Jaeyeon, Mythili, and Mike have helped me
refine my research ideas.  Sheila Marian made my life around the lab
smooth and pleasant; I will miss discussing the latest Red Sox gossip
with her.  I also thank the members of the NMS and PDOS groups for
listening to my many thoughts and practice talks.

I am fortunate to have wonderful colleagues outside of MIT.  In
particular, I would like to thank Randy Bush, Tim Griffin, Albert
Greenberg, Jacobus van der Merwe, Aman Shaikh, Lixin Gao,
kc claffy, Avi Freedman, Morley Mao, Olivier Bonaventure, Richard
Mortier, and the network operators who have used \rccns, for
taking a deep interest in my work and providing feedback and support.  I
thank Joan Feigenbaum and Anthony Joseph for their advice and insights
during my job search.  I've had helpful discussions (and great times)
with Aditya Akella, Anukool Lakhina, Ratul Mahajan, Joel Sommers,
Lakshmi Subramanian, and Renata Teixeira.  Susie Wee and John
Apostolopoulos introduced me to the rewards of research and have
offered useful professional advice.

Mike Freedman, Sean Montgomery, and Ajay Kulkarni have continually
supported me for nearly ten years.  Anukool Lakhina has been an
indispensable friend during the writing of this dissertation.  Greg
Harfst, Claire Monteleoni, Jamie Fine, Jennifer Bowen, Jonathan Hall,
Jeremy Stribling,
Kelly Carleton, Carson Reynolds, and Steven Richman have been companions
in both celebration and commiseration.  Dave Andersen, Thomer Gil, Alex
Yip, Sam Madden, Dye-Zone Chen, and the members of the MIT Cycling
Club were all great cycling partners.  Sanmay Das, Tony Ezzat,
and Jos\'{e} Rafael Galvanis gave me many memories on the squash courts,
and Doug DeCouto made swimming less dull.

The research in this dissertation was funded through an NSF Graduate
Research Fellowship, the NSF under Cooperative Agreement ANI-0225660,
the Defense Advanced Research Projects Agency (DARPA) and the Space and
Naval Warfare Systems Center, San Diego, under contract
N66001-00-1-8933, and a Cisco URP grant.  The writing of this
dissertation was fueled with sustenance from the 1369 Coffeehouse and
Mariposa Caf\'{e} in Cambridge, MA; and the Big Cup, the Saurin Park
Caf\'{e}, and Grounded in New York City.

My parents, Carolyn and Scott Feamster, 
have always encouraged me to push the frontier of
knowledge.  Both my 
parents and my late grandmother, Elizabeth Huey, have been a constant
source of love and inspiration, without which this accomplishment would
have been impossible.  In return, I dedicate this dissertation to them.
My success is also theirs.


%\vskip 0.3in
%\noindent\textbf{\em Bibliographic Notes}

\chapter*{Bibliographic Notes}
\addcontentsline{toc}{chapter}{Bibliographic Notes}

\noindent
An early version of some material from Chapter~\ref{chap:rlogic} appears
in a paper co-authored with Hari Balakrishnan~\cite{Feamster2003b}.
Material from Chapter~\ref{chap:rcc} appears in a paper co-authored with
Hari Balakrishnan~\cite{Feamster2004h}.  Material from
Chapter~\ref{chap:sandbox} appears in papers co-authored with Jennifer
Rexford and Jared Winick~\cite{Feamster2002b, Feamster2004}.  Material
from Chapter~\ref{chap:policy} appears in a paper co-authored with
Ramesh Johari and Hari Balakrishnan~\cite{Feamster2005b}.  
The original design for the Routing Control Platform, which is briefly
discussed in Chapter~\ref{chap:concl}, appears in a paper co-authored
with Hari Balakrishnan, Jennifer Rexford, Aman Shaikh, and Jacobus van
der Merwe~\cite{feamster:fdna2004}.  Ramesh Johari's dissertation
provided formatting and typography inspiration.


%The material
%from Section~\ref{sec:dynamic} dealing with detection of inconsistent
%route advertisements appears in a paper co-authored with Zhuoqing Morley
%Mao and Jennifer Rexford~\cite{Feamster2004b}; Morley performed the
%evaluation of these detection algorithms as presented in
%Section~\ref{subsec:ibgpResult}.  



